---
title: "BIOST 578 Homework 1"
author: "Marea Cobb"
date: "January 15, 2015"
output: html_document
---

Reviewer: Gloria Chi

Loads specific libraries 


```{r}
library(ggplot2)
library(RColorBrewer)
library(reshape2)
```

Loads the iris data


```{r}
data(iris)
head(iris)
```

Creates a new column for flower_id for each entry.


```{r}
iris$flower_id <- rownames(iris)
head(iris)
```

Uses the default melt method to rearrange the data for easier classification.


```{r}
iris_melted <- melt(iris)
```


```{r}
## Using Species, flower_id as id variables
head(iris)
```

Splits the variable at the "." to get the value name (length/width and flower type)


```{r}
split_variable <- strsplit(as.character(iris_melted$variable), split="\\.")
```

Creates the variable "flower part" from the first value in split_variable and adds it to the table


```{r}
iris_melted$flower_part <- sapply(split_variable, "[", 1)
```

Creates the variable "measurement type" from the second value in split_variable and adds it to the table


```{r}
iris_melted$measurement_type <- sapply(split_variable, "[", 2)
```

Removes variable from the table as we no longer need it


```{r}
iris_melted$variable <- NULL
```

Reshapes the data so width and length are now separte columns in the table 


```{r}
iris_cast <- dcast(iris_melted, formula=flower_id+Species+flower_part~measurement_type)
head(iris_cast)
```

Creates a plot displaying the length and width of petals and sepals classified by flower type.


```{r}
ggplot(data=iris_cast, colour=colors, aes(x=Width, y=Length, color=Species))+ # Add points and use free scales in the facet
  geom_point()+facet_grid(Species~flower_part, scale="free")+
  #' Adds a regression line
  geom_smooth(method="lm")+
  theme(plot.background = element_rect(colour="blue"))
```

